---
title: "Machine Learning-Clustering"
description: |
  An example of a machine learning model for dimensionality reduction
author:
  - name: Karis Cox- from the class of Dr. Hunt
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(caret)
trainIndex <- createDataPartition(iris$Species, p = .6, list = FALSE, times = 1)

irisTrain <- iris[ trainIndex,]
irisTest  <- iris[-trainIndex,]

preProcValues <- preProcess(irisTrain, method = c("center", "scale"))
trainTransformed <- predict(preProcValues, irisTrain)

preProcValues <- preProcess(irisTest, method = c("center", "scale"))
testTransformed <- predict(preProcValues, irisTest)
```
#Clustering

In the following example I use spacial clustering to identify which data points belong to each classification. I use a k-means algorithm, one of the simplest to understand. I use 2, 3, and 4 clusters to see how well the data is centered around these cluster centers. 

This is the full data set based on 2 clusters in a table form. 
```{r}
TwoClusters<-kmeans(trainTransformed[,-5],centers=2)
Clusterdata<-trainTransformed
Clusterdata$Cluster<-as.factor(TwoClusters$cluster)
knitr::kable(Clusterdata)%>%
  kableExtra::kable_styling("striped")%>%
  kableExtra::scroll_box(width = "100%",height="300px")
```

Here are a few graphic representations showing how the data is split when 2 clusters are used. 

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)+facet_wrap(~Species)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Species))+geom_point(alpha=0.5) + 
  geom_point(data=as.data.frame(TwoClusters$centers), aes(color="Cluster center"), size=5) + theme(legend.title = element_blank())+ggtitle("Iris Cluster Demonstration")
```

This is a table of the data when 3 data clusters are used. 

```{r}
ThreeClusters<-kmeans(trainTransformed[,-5],centers=3)
Clusterdata<-trainTransformed
Clusterdata$Cluster<-as.factor(ThreeClusters$cluster)
knitr::kable(Clusterdata)%>%
  kableExtra::kable_styling("striped")%>%
  kableExtra::scroll_box(width = "100%",height="300px")
```

These graphs show how the data is split when 3 clusters are used rather than 2. As you can see, using 3 centers allows for a more accurate split of the data. 

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)+facet_wrap(~Species)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Species))+geom_point(alpha=0.5) + 
  geom_point(data=as.data.frame(ThreeClusters$centers), aes(color="Cluster center"), size=5) + theme(legend.title = element_blank())+ggtitle("Iris Cluster Demonstration")
```

And lastly, here is a table form of the data when 4 clusters are used. 

```{r}
FourClusters<-kmeans(trainTransformed[,-5],centers=4)
Clusterdata<-trainTransformed
Clusterdata$Cluster<-as.factor(FourClusters$cluster)
knitr::kable(Clusterdata)%>%
  kableExtra::kable_styling("striped")%>%
  kableExtra::scroll_box(width = "100%",height="300px")
```

Here are some visual representations of what the data would look like when split between 4 cluster centers. Because there are only 3 species, 4 cluster centers does not prove to be extra beneficial. 

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Cluster))+geom_point(alpha=0.5)+facet_wrap(~Species)
```

```{r}
ggplot(data=Clusterdata,mapping = aes(x=Sepal.Width,y=Petal.Width,color=Species))+geom_point(alpha=0.5) + 
  geom_point(data=as.data.frame(FourClusters$centers), aes(color="Cluster center"), size=5) + theme(legend.title = element_blank())+ggtitle("Iris Cluster Demonstration")
```



